---
title: "A Tale of Two Cities"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    self_contained: false
    theme: spacelab
    
---

Summary {.storyboard}
=========================================

### Summary

```{r, echo=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(readxl))
suppressMessages(library(tinytex))
suppressMessages(library(plotly))
suppressMessages(library(lubridate))
suppressMessages(library(htmltools))
suppressMessages(library(flexdashboard))
Football.Revenues <- read.csv(file="footiedata/FootballRevenues.csv", header=TRUE, sep=",")
## This creates an object (in this case a data.frame) named 'Football.Revenues' from the csv file.
LFCvMCFC <- read_xlsx("footiedata/LFCVMCFC.xlsx")
footie_palette <- c(rgb(200,16,46, maxColorValue = 255), rgb(108,171,221, maxColorValue = 255))
footie_palette <- setNames(footie_palette, c("LFC", "MCFC"))
LabelTags <- c("Liverpool FC", "Manchester City")
LabelTags <- setNames(LabelTags, c("LFC", "MCFC"))
LFCvMCFC.Revenues <- # assignment to an object.
 filter(Football.Revenues, Identifier %in% c("LFC","MCFC")) # we like to use %in% but there are other methods.
Revenue.Totals <-
  LFCvMCFC.Revenues %>% # we will discuss the "pipe" %>% below.
  group_by(Club) %>%
    summarize(Total = sum(Total))
SpendSummary <-
  LFCvMCFC %>%
  group_by(Club) %>%
  summarise(TotalFees = sum(Fee))

SummarySub <- plot_ly()%>%
  add_pie(data = Revenue.Totals, labels = ~Club, values = ~Total,
    marker = list(colors = footie_palette),
    name = "Total Revenue",
    domain = list(row = 0, column = 0),
    textinfo = 'label+value+percent')%>%
  layout(annotations = list(x = 0.15 , y = .25, text = "Total Revenue", showarrow = F, 
xref='paper', yref='paper'))%>%
  
  
  add_pie(data = SpendSummary, labels = LabelTags, values = ~TotalFees,
    marker = list(colors = footie_palette),
    name = "Transfer Spending",
    domain = list(row = 0, column = 1),
    textinfo = 'label+value+percent')%>%
  
layout(annotations = list(x = 0.85 , y = .25, text = "Transfer Spend", showarrow = F, 
xref='paper', yref='paper'),
title = "Liverpool and Manchester City - Revenue and Transfer Spend 2008 to 2018
       (£ Millions)", showlegend = F,
         grid=list(rows=1, columns=2),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

SummarySub
```

***
As we can see, from season 2007/2008 to season 2017/2018, revenues show only a minimal advantage for Manchester City. However, as we shall see with some upcoming visualizations, the trend would indicate a growing advantage for Manchester City.

On the other hand, transfer spending has been heavily tilted in favour of Manchester City since season 2007/2008.

A note on sources:
Revenue data is based on Deloitte Football Money League publications and can be viewed [HERE](https://www2.deloitte.com/uk/en/pages/sports-business-group/articles/deloitte-football-money-league.html)

Transfer data is collated from multiple sources; primarily https://www.transfermarkt.com/ and https://www.transferleague.co.uk/.

A good deal of inspiration that started us on this project came from https://github.com/ewenme who did a fine job scraping transfermarkt. We will definitely be going back to that dataset for future analysis. His football repo can be viewed at https://github.com/ewenme/footy-transfer-data

Revenue {.storyboard}
=========================================

### Total Annual Revenue

```{r, echo=FALSE}
Revenue.Plot <-
  
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Total,
    type = "scatter",
    mode = "line",
    color = ~Identifier,
    colors = footie_palette) %>%
  
  layout(legend = list(x = 0.05, y = 0.95), #sometimes we move the legend around
      title = "LFC & MCFC Annual Revenue",
         yaxis = list(title = '(£ Millions)'))

Revenue.Plot
```

### Annual Revenue by Stream

```{r}
Matchday.Plot <-
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Matchday,
    type = 'scatter',
    mode = 'lines',
    color = ~Identifier,
    colors = footie_palette,
    showlegend = FALSE
  ) %>%
  
  layout(annotations = list(x = 0.2 , y = .85, text = "Matchday", showarrow = F, 
xref='paper', yref='paper'))

Broadcast.Plot <-
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Broadcast,
    type = 'scatter',
    mode = 'lines',
    color = ~Identifier,
    colors = footie_palette,
    showlegend = FALSE
  ) %>%

layout(annotations = list(x = 0.4 , y = .85, text = "Broadcast", showarrow = F, 
xref='paper', yref='paper'))
  
Commercial.Plot <-
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Commercial,
    type = 'scatter',
    mode = 'lines',
    color = ~Identifier,
    colors = footie_palette
  ) %>%
  
  
layout(legend = list(x = 0.04, y = 0.99),
  annotations = list(x = 0.6 , y = .85, text = "Commercial", showarrow = F, 
xref='paper', yref='paper'))

Stream.Plot <- 
  subplot(Matchday.Plot, Broadcast.Plot, Commercial.Plot, shareY = TRUE) %>%
 
layout(title = "LFC & MCFC Revenue Categories",
      yaxis = list(title = '(£ Millions)')
)

Stream.Plot
```

Transfers {.storyboard}
=========================================
