---
title: "A Tale of Two Cities"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    self_contained: false
    theme: spacelab
    
---

Summary {.storyboard}
=========================================

### Summary

```{r, echo=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(readxl))
suppressMessages(library(tinytex))
suppressMessages(library(plotly))
suppressMessages(library(lubridate))
suppressMessages(library(htmltools))
suppressMessages(library(flexdashboard))
Football.Revenues <- read.csv(file="data/FootballRevenues.csv", header=TRUE, sep=",")
## This creates an object (in this case a data.frame) named 'Football.Revenues' from the csv file.
LFCvMCFC <- read_xlsx("data/LFCVMCFC.xlsx")
footie_palette <- c(rgb(200,16,46, maxColorValue = 255), rgb(108,171,221, maxColorValue = 255))
footie_palette <- setNames(footie_palette, c("LFC", "MCFC"))
LabelTags <- c("Liverpool FC", "Manchester City")
LabelTags <- setNames(LabelTags, c("LFC", "MCFC"))
LFCvMCFC.Revenues <- # assignment to an object.
 filter(Football.Revenues, Identifier %in% c("LFC","MCFC")) # we like to use %in% but there are other methods.
Revenue.Totals <-
  LFCvMCFC.Revenues %>% # we will discuss the "pipe" %>% below.
  group_by(Club) %>%
    summarize(Total = sum(Total))
SpendSummary <-
  LFCvMCFC %>%
  group_by(Club) %>%
  summarise(TotalFees = sum(Fee))

SummarySub <- plot_ly()%>%
  add_pie(data = Revenue.Totals, labels = ~Club, values = ~Total,
    marker = list(colors = footie_palette),
    name = "Total Revenue",
    domain = list(row = 0, column = 0),
    textinfo = 'label+value+percent')%>%
  layout(annotations = list(x = 0.15 , y = .25, text = "Total Revenue", showarrow = F, 
xref='paper', yref='paper'))%>%
  
  
  add_pie(data = SpendSummary, labels = LabelTags, values = ~TotalFees,
    marker = list(colors = footie_palette),
    name = "Transfer Spending",
    domain = list(row = 0, column = 1),
    textinfo = 'label+value+percent')%>%
  
layout(annotations = list(x = 0.85 , y = .25, text = "Transfer Spend", showarrow = F, 
xref='paper', yref='paper'),
title = "Liverpool and Manchester City - Revenue and Transfer Spend 2008 to 2018
       (£ Millions)", showlegend = F,
         grid=list(rows=1, columns=2),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

SummarySub
```

***
As we can see, from season 2007/2008 to season 2017/2018, revenues show only a minimal advantage for Manchester City. However, as we shall see with some upcoming visualizations, the trend would indicate a growing advantage for Manchester City.

On the other hand, transfer spending has been heavily tilted in favour of Manchester City since season 2007/2008.

A note on sources:
Revenue data is based on Deloitte Football Money League publications and can be viewed [HERE](https://www2.deloitte.com/uk/en/pages/sports-business-group/articles/deloitte-football-money-league.html)

Transfer data is collated from multiple sources; primarily https://www.transfermarkt.com/ and https://www.transferleague.co.uk/.

A good deal of inspiration that started us on this project came from https://github.com/ewenme who did a fine job scraping transfermarkt. We will definitely be going back to that dataset for future analysis. His football repo can be viewed at https://github.com/ewenme/footy-transfer-data

Revenue {.storyboard}
=========================================

### Total Annual Revenue

```{r, echo=FALSE}
Revenue.Plot <-
  
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Total,
    type = "scatter",
    mode = "line",
    color = ~Identifier,
    colors = footie_palette) %>%
  
  layout(legend = list(x = 0.05, y = 0.95), #sometimes we move the legend around
      title = "LFC & MCFC Annual Revenue",
         yaxis = list(title = '(£ Millions)'))

Revenue.Plot
```

***
Liverpool booked more revenue than Manchester City from the 2007/2008 season until 2010/2011 season. From 2011/2012 onwards, Manchester City has steadily outperformed Liverpool. Although, for 2017/2018 it appears Liverpool have shrunk the gap somewhat. Whether this becomes a future trend is impossible to say but worth watching. In any event, it is clear that over 11 seasons, the overall revenue for these two teams had a steep incline. No doubt this is a large reflection of the tv deals signed by the Premier League over the past decade and the continued expansion of Premier League commercial activity.

### Annual Revenue by Stream

```{r}
Matchday.Plot <-
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Matchday,
    type = 'scatter',
    mode = 'lines',
    color = ~Identifier,
    colors = footie_palette,
    showlegend = FALSE
  ) %>%
  
  layout(annotations = list(x = 0.2 , y = .85, text = "Matchday", showarrow = F, 
xref='paper', yref='paper'))

Broadcast.Plot <-
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Broadcast,
    type = 'scatter',
    mode = 'lines',
    color = ~Identifier,
    colors = footie_palette,
    showlegend = FALSE
  ) %>%

layout(annotations = list(x = 0.4 , y = .85, text = "Broadcast", showarrow = F, 
xref='paper', yref='paper'))
  
Commercial.Plot <-
  plot_ly(
    data = LFCvMCFC.Revenues,
    x = ~Season,
    y = ~Commercial,
    type = 'scatter',
    mode = 'lines',
    color = ~Identifier,
    colors = footie_palette
  ) %>%
  
  
layout(legend = list(x = 0.04, y = 0.99),
  annotations = list(x = 0.6 , y = .85, text = "Commercial", showarrow = F, size = 14,
xref='paper', yref='paper'))

Stream.Plot <- 
  subplot(Matchday.Plot, Broadcast.Plot, Commercial.Plot, shareY = TRUE) %>%
 
layout(title = "LFC & MCFC Revenue Categories",
      yaxis = list(title = '(£ Millions)')
)

Stream.Plot
```

***
This certainly tells a story.

1. Matchday revenue is less relied upon than broadcast or commercial. After some remodelling at Anfield, Liverpool has opened an advantage.
2. Broadcast revenue has seen huge growth and jumps can be seen every 3 years as new broadcast contracts are negotiated. Both teams are relatively even here with some fluctuation based on Champions League participation and other factors.
3. Commercial revenue is significantly disparate. The latest figures show Manchester City a whopping £84 Million ahead for 2017/2018. While Liverpool is also growing its commercial revenue, Manchester City have held this advantage for some time now.

Deloitte data does not include player sales in revenues. Commercial revenue is typically made up of sponsorship deals, leasing venues for concerts, etc. It is interesting that Manchester City is able to grow this type of activity beyond the levels of Liverpool.

Transfers {.storyboard}
=========================================

### Annual Transfer Spending

```{r}
#add column to df extracting year from "Date"
LFCvMCFC$Year <-
year(LFCvMCFC$Date)

AnnualSpending <-
  LFCvMCFC %>%
  group_by(Year, Club) %>%
  summarise(AnnualSpend = sum(Fee))

AnnualPlot <-
  
  plot_ly(
    data = AnnualSpending,
    x = ~Year,
    y = ~AnnualSpend,
    type = "scatter",
    mode = "line",
    color = ~Club,
    colors = footie_palette) %>%
  
  layout(legend = list(x = 0.05, y = 0.95), #sometimes we move the legend around
      title = "LFC & MCFC Annual Transfer Spending",
         yaxis = list(title = 'Spent (£ Millions)'))

AnnualPlot
```

***
Manchester City took full advantage of new investment in 2008, 2009 and 2010. This pattern also resumed in 2015, 2016 and 2017 where significant investment was made. Liverpool spent more on transfers than Manchester City during only three years; 2011, 2014 and 2018.

### Individual Transfers

```{r}
TransactionScatter <- 
 
   plot_ly(
    data = LFCvMCFC,
    x = ~Year,
    y = ~Fee,
    type = "scatter",
    mode = 'markers',
    color = ~Club,
    colors = footie_palette,
    size = 10, #we have opted for a standard size here
    marker = list(opacity = 0.75), #we have a lot of density in parts so opacity is increased
    text = ~paste(Player_Purchased, Fee, Club), #text and hoverinfo introduced to improve default hovertext
    hoverinfo = "text") %>%

layout(legend = list(x = 0.06, y = 1.0),
       title = "LFC & MCFC Individual Transfers",
       yaxis = list(title = 'Cost (£ Millions)'))

TransactionScatter
```

***
We can see LFC has the highest amount paid for a single player (Virgil van Dijk in 2018) but in all other years, MCFC purchased the most expensive player; in many years, several more expensive players than LFC.

### Age Distribution of Players Purchased

```{r}
AgeViola <- LFCvMCFC %>%
  
  plot_ly(
    x = ~Club,
    y = ~Age_at_Transfer,
    split = ~Club,
    type = 'violin',
    colors = footie_palette,
    color = ~Club,
    box = list(visible = T),
    meanline = list(visible = T)) %>% 
  
  layout(title = "LFC & MCFC Age Distribution of Transfer Targets",
    xaxis = list(title = "Club"),
        yaxis = list(title = "Age at Transfer",
      zeroline = F)
  )

AgeViola
```

***
Liverpool demonstrates a mean age of transfer targets almost a full year younger than Manchester City. This is also true of the q3 values and slightly less true of q1. The difference is reduced somewhat looking at median; KDE of 1 occurs for LFC at ~24.5 years of age and for MCFC at ~24.80. Again, readers can reach their own conclusions but we find this interesting and would be of even greater interest if combined with data regarding contract length, etc
