---
title: "A Tale of Two Cities"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    self_contained: false
---

Visual Storyboard {.storyboard}
=========================================

### Summary

```{r}

SummarySub <- plot_ly()%>%
  add_pie(data = Revenue.Totals, labels = ~Club, values = ~Total,
    marker = list(colors = footie_palette),
    name = "Total Revenue",
    domain = list(row = 0, column = 0),
    textinfo = 'label+value+percent')%>%
  layout(annotations = list(x = 0.15 , y = .25, text = "Total Revenue", showarrow = F, 
xref='paper', yref='paper'))%>%
  
  
  add_pie(data = SpendSummary, labels = LabelTags, values = ~TotalFees,
    marker = list(colors = footie_palette),
    name = "Transfer Spending",
    domain = list(row = 0, column = 1),
    textinfo = 'label+value+percent')%>%
  
layout(annotations = list(x = 0.85 , y = .25, text = "Transfer Spend", showarrow = F, 
xref='paper', yref='paper'),
title = "Liverpool and Manchester City - Revenue and Transfer Spend 2008 to 2018
       (£ Millions)", showlegend = F,
         grid=list(rows=1, columns=2),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))


SummarySub
```

***
As we can see, from season 2007/2008 to season 2017/2018, revenues show only a minimal advantage for Manchester City. However, as we shall see with some upcoming visualizations, the trend would indicate a growing advantage for Manchester City.

On the other hand, transfer spending has been heavily tilted in favour of Manchester City since season 2007/2008.

A note on sources:
Revenue data is based on Deloitte Football Money League publications and can be viewed [HERE](https://www2.deloitte.com/uk/en/pages/sports-business-group/articles/deloitte-football-money-league.html)

Transfer data is collated from multiple sources; primarily https://www.transfermarkt.com/ and https://www.transferleague.co.uk/.

A good deal of inspiration that started us on this project came from https://github.com/ewenme who did a fine job scraping transfermarkt. We will definitely be going back to that dataset for future analysis. His football repo can be viewed at https://github.com/ewenme/footy-transfer-data

### Revenues

```{r}
```

### Transfers

Details
=========================================

Column
-----------------------------------------
```

